---
title: '`Course Project` Statistical Inference'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  author: "Anderson H Uyekita"
  course: 'Statistical Inference'
  course_url: 'https://www.coursera.org/learn/statistical-inference'
  specialization: 'Data Science: Statistics and Machine Learning Specialization'
  specialization_url: 'https://www.coursera.org/specializations/data-science-statistics-machine-learning'
  instructor: "Brian Caffo"
  course_start: !r lubridate::ymd("2022/07/01", tz = "America/Sao_Paulo")
  course_finish: !r lubridate::today(tzone = "America/Sao_Paulo")
  week: "Week 4"
  slug: 'Week%204'
  rpubs: 'https://rpubs.com/AndersonUyekita/course-project_statistical-inference'
  project_repo: 'https://github.com/AndersonUyekita/statistical-inference_course-project'
  codebook: 'https://github.com/AndersonUyekita/statistical-inference_course-project/blob/master/codebook.md'
  instructions: 'https://github.com/AndersonUyekita/statistical-inference_course-project/blob/master/instructions.md'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo=TRUE,cache=TRUE,message=FALSE,warning=FALSE)
library(rmarkdown)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`

--------------------------------------------------------------------------------

## R2D2 C3PO

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/AndersonUyekita/statistical-inference_course-project/master?urlpath=rstudio)

### Synopsis


--------------------------------------------------------------------------------

## 1. Introduction

The project consists of two parts:

* A simulation exercise.
* Basic inferential data analysis.

## 2. Objectivies [](){id="objectivies"}


## 3. Requirements and Settings


```{r libraries,echo=TRUE,warning=FALSE,message=FALSE,include=TRUE}
# Loading libraries
library(rmarkdown)
library(tidyverse)
library(ggplot2)
library(magrittr)
```


```{r language,echo=TRUE,warning=FALSE,message=FALSE,include=TRUE,cache=TRUE}
# Force the R environment to use English settings
Sys.setlocale("LC_ALL","English")
```


```{r session-info,echo=FALSE}
sessionInfo()
```

## 4. Data Processing



### 4.1. Downloading

Creating a data folder to store the dataset from NOAA.



### 4.2. Loading



## 5. Data Assessing



### 5.1. Dimensions


### 5.2. Head and Tail


### 5.3. Variables


## 6. Data Manipulation


### 6.1. Subsetting the raw data


### 6.2. Merging Damage and Exponential Variables


### 6.3. Uniforming Event Types


### 6.3. Tidy dataset

## 7. Results [](){id="results"}
